Hrvatska u kontekstu europskih konvergencijskih klubova

### 2024-08-28 Paper.Rmd
los

```{r libs,  echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#### LIBS ----
library(eurostat)
library(tidyverse)
library(zoo)
library(scales)
library(tseries)
library(urca)
library(fracdiff)
library(arfima)
library(gt)

@@

```{r ljbx,  echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
compute_ljung_box_every_second_lag <- function(data, max_lag = 20) {
  results <- list()
  for (lag in seq(2, max_lag, by = 2)) {  # Loop through every second lag
    p_value <- Box.test(data$diff_log_gdp_per_capita, lag = lag, type = "Ljung-Box")$p.value
    results[[lag]] <- tibble(lag = lag, p_value = p_value)
  }
  bind_rows(results) %>% mutate(group = data$group[1])
}

@@

# Display the results as a table
print(ljung_box_table_second_lags)

# Create a nice table visualization using gt
ljung_box_gt_second_lags <- ljung_box_table_second_lags %>%
  gt() %>%
  tab_header(
    title = "Ljung-Box Test P-values for Every Second Lag",
    subtitle = "P-values for every second lag up to 20 for each group"
  ) %>%
  cols_label(
    group = "Group",
    lag_2 = "Lag 2", lag_4 = "Lag 4", lag_6 = "Lag 6", lag_8 = "Lag 8", lag_10 = "Lag 10",
    lag_12 = "Lag 12", lag_14 = "Lag 14", lag_16 = "Lag 16", lag_18 = "Lag 18", lag_20 = "Lag 20"
  ) %>%
  fmt_number(
    columns = starts_with("lag_"),
    decimals = 3
  ) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 14,
    heading.subtitle.font.size = 12,
    column_labels.font.size = 12
  )

# Print the gt table
#print(ljung_box_gt_second_lags)
```

@@

```{r acf,  echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

@@

```{r adf, libs,  echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}
# Calculate the differences between Croatia's log GDP per capita and the groups

@@

# Display the ADF test results table
print(adf_results)
```

@@

```{r frac, eval=FALSE, echo=FALSE, message=FALSE, warning=FALSE, fig.width=10, fig.height=6}

@@

@@

# P-GPH estimator
pgph_test_results <- function(data, group_name) {
  result <- urca::ur.pp(data$diff_log_gdp_per_capita, type = "Z-alpha", model = "trend")

}

pgph_croatia_vs_nms8 <- pgph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8")
pgph_croatia_vs_nms12 <- pgph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12")
pgph_croatia_vs_se4 <- pgph_test_results(diff_croatia_vs_se4, "Croatia vs SE4")
pgph_croatia_vs_eu15 <- pgph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15")

@@

# Combine all results into a single data frame for easy comparison
all_results <- bind_rows(
  robinson_results %>% mutate(test = "Robinson"),
  gph_results %>% mutate(test = "GPH"),
  pgph_results %>% mutate(test = "P-GPH")
)

# Print all results
print(all_results)
```

>>>>

```{r libs,  echo=FALSE, message=FALSE, warning=FALSE, results='hide'}
#### LIBS ----
library(eurostat)
library(tidyverse)
library(zoo)
library(scales)
library(tseries)
library(urca)
library(fracdiff)
library(arfima)
library(gt)
library(kableExtra)
library(urca)

@@

```{r ljbx,  echo=FALSE, message=FALSE, warning=FALSE,results='asis', fig.width=10, fig.height=6}
compute_ljung_box_every_second_lag <- function(data, max_lag = 20) {
  results <- list()
  for (lag in seq(2, max_lag, by = 2)) {  # Loop through every second lag
    p_value <- Box.test(data$diff_log_gdp_per_capita, lag = lag, type = "Ljung-Box")$p.value
    results[[lag]] <- tibble(lag = lag, p_value = p_value)
  }
  bind_rows(results) %>% mutate(group = data$group[1])
}

@@

# Display the results as a table
#print(ljung_box_table_second_lags)

# Create a nice table visualization 
ljung_box_kable <- ljung_box_table_second_lags %>%
  kable(
    caption = "Ljung-Box Test P-values for Every Second Lag",
    col.names = c("Group", "Lag 2", "Lag 4", "Lag 6", "Lag 8", "Lag 10",
                  "Lag 12", "Lag 14", "Lag 16", "Lag 18", "Lag 20"),
    format = "html",
    digits = 3,
    align = "c"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#D3D3D3") %>%
  add_header_above(c(" " = 1, "Ljung-Box Test P-values" = 10))

# Print the kable table
print(ljung_box_kable)
```

@@

```{r acf,  echo=FALSE, message=FALSE, warning=FALSE,results='asis', fig.width=10, fig.height=6}

@@

```{r adf,  echo=FALSE, message=FALSE, warning=FALSE,results='asis', fig.width=10, fig.height=6}
# Calculate the differences between Croatia's log GDP per capita and the groups

@@

# Display the ADF test results table
#rint(adf_results)

# Function to add stars based on significance
add_significance_stars <- function(value) {
  rounded_value <- round(value, 4)  # Round to 4 decimal places
  if (value < -3.43) {
    return(paste0(rounded_value, "***"))
  } else if (value < -2.86) {
    return(paste0(rounded_value, "**"))
  } else if (value < -2.57) {
    return(paste0(rounded_value, "*"))
  } else {
    return(as.character(rounded_value))
  }
}

# Apply the significance function to each column
adf_results_stars <- adf_results %>%
  mutate(
    `Bez kons/tr` = sapply(`Bez kons/tr`, add_significance_stars),
    Konst = sapply(Konst, add_significance_stars),
    `Konst/tr` = sapply(`Konst/tr`, add_significance_stars)
  )

# Create the kable table with significance stars
adf_results_kable <- adf_results_stars %>%
  select(group_name, `Bez kons/tr`, Konst, `Konst/tr`) %>%
  kable(
    caption = "ADF Test Results with Significance",
    col.names = c("Group Name", "Bez kons/tr", "Konst", "Konst/tr"),
    format = "html",
    align = "c"
  ) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), full_width = F) %>%
  column_spec(1, bold = TRUE) %>%
  row_spec(0, bold = TRUE, background = "#D3D3D3")

# Print the table
print(adf_results_kable)

@@

```{r fracr1 ,  echo=FALSE, message=FALSE, warning=FALSE,  fig.width=10, fig.height=6}

@@

```{r fracr2 ,  echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, fig.height=6}

@@

```{r fracr3 ,  echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.width=10, fig.height=6}
# Define the P-GPH estimator function
pgph_test_results <- function(data, group_name) {
  # Run the Phillips-Perron test (as a proxy for P-GPH)
  result <- ur.pp(data$diff_log_gdp_per_capita, type = "Z-alpha", model = "trend")

  # Extract the test statistic
  test_stat <- result@teststat[1]  # Z-alpha statistic

  # Check the structure of the cval object
  critical_values <- result@cval

  # Safely extract critical values
  critical_1_percent <- ifelse(nrow(critical_values) >= 1, critical_values[1, 1], NA)
  critical_5_percent <- ifelse(nrow(critical_values) >= 2, critical_values[2, 1], NA)
  critical_10_percent <- ifelse(nrow(critical_values) >= 3, critical_values[3, 1], NA)

  # Return as a data frame (or tibble)
  tibble(
    group_name = group_name,
    pgph_stat = test_stat,
    critical_1_percent = critical_1_percent,
    critical_5_percent = critical_5_percent,
    critical_10_percent = critical_10_percent
  )
}

# Apply the P-GPH estimator to each group
pgph_croatia_vs_nms8 <- pgph_test_results(diff_croatia_vs_nms8, "Croatia vs NMS8")
pgph_croatia_vs_nms12 <- pgph_test_results(diff_croatia_vs_nms12, "Croatia vs NMS12")
pgph_croatia_vs_se4 <- pgph_test_results(diff_croatia_vs_se4, "Croatia vs SE4")
pgph_croatia_vs_eu15 <- pgph_test_results(diff_croatia_vs_eu15, "Croatia vs EU15")

@@
